---
title: "Horror Movies"
author: "Fruzsina Hejja"
date: "2024-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(car)
library(lubridate)
```

In the final project, I will use a dataset containing horror movies and information based on their production and popularity, likeability.

##Reading the dataset
```{r}
horror_movies <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2022/2022-11-01/horror_movies.csv") #Read from the TidyTuesday github page.
```
##Checking the structure of the dataset
```{r}
str(horror_movies)
```
#Exploratory data analysis
Firs a quick overview of all the variables
```{r}
summary(horror_movies)
```
##Creating and exploring the year variable
```{r}
horror_movies <- horror_movies %>%

  mutate(year = year(release_date))
```
###Checking which year most of the movies are from in the dataset
```{r}
horror_movies %>%
  count(year) %>%
  arrange(desc(n))

ggplot(horror_movies, aes(x = year)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "white") +
  labs(title = "Number of horror movies per year",
       x = "Year",
       y = "Number of movies")
  theme_minimal()
```
##Filtering out movies less voted for and exploring average ratings
```{r}
horror_movies <- horror_movies %>%
  filter(vote_count > 1000)

summary(horror_movies$vote_count)


ggplot(horror_movies, aes(x = vote_average)) +
  geom_histogram(bins = 20, fill = "skyblue", color = "white") +
  labs(title = "Distribution of vote averages for horror movies",
       x = "Vote average",
       y = "Number of movies") +
  theme_minimal()
```
###What are the top 10 movies based on the average vote?
```{r}
horror_movies %>%
  arrange(desc(vote_average)) %>%
  select(title, vote_average) %>%
  head(10)
```
##What are the co-occuring genres in horror movies?
```{r}
horror_movies %>% #What is the most common co-oocuring genre with horror movies
  separate_rows(genre_names, sep = ", ") %>%
  rename(genre = genre_names) %>%
  filter(genre != "Horror") %>%
  count(genre) %>%
  ggplot(aes(x = n, y = fct_reorder(genre, n))) +
  geom_col(fill = "slateblue") +
  labs(title = "Most common co-occuring genres with horror movies",
       x = "Number of movies",
       y = "Genre") +
  theme_minimal()
```
##Exploring budget and revenue
###Removing missing data from the budget and revenue variables
```{r}
horror_movies <- horror_movies %>%
  filter(budget > 0, revenue > 0)
```
##Checking budget vs revenue with scatterplot
```{r}
ggplot(horror_movies, aes(x = budget, y = revenue)) +
  geom_point(color = "skyblue") +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::dollar) +
  labs(title = "Budget and revenue of horror movies",
       x = "Budget",
       y = "Revenue") +
  theme_minimal()
  
```
###What movie was the most expensive to make?
```{r}
horror_movies %>%
  arrange(desc(budget)) %>%
  select(title, budget) %>%
  head(1)
```
###What movie had the highest revenue?
```{r}
horror_movies %>%
  arrange(desc(revenue)) %>%
  select(title, revenue) %>%
  head(1)
```
###How much did that movie cost?
```{r}
horror_movies %>%
  filter(title == "It") %>%
  select(title, budget)
```
#Model building
Assuming that a higher budget results in a higher quality movie that could be more enjoyable for the audiance, the average vote of the horror movies could be predicted based on budget.
```{r}
modelA <- lm(vote_average ~ budget,
             data = horror_movies)
summary(modelA)
```

The results show, that the regression model is statistically significant (*F = 4.967*, *p < 0.05*), but it explains only 1.25% of the variance (*R^2 = 0.0125*).

##What other predictors could explain the averahe vote of a horror movie?
```{r}
modelB <- lm(vote_average ~ budget +
               vote_count +
               revenue,
             data = horror_movies)
summary(modelB)
```
The results show, that the more complex model using multiple predictors is statistically significant (*F = 42.93*, *p < 0.01*), and it explains 24.7% of the variance (*R^2 = 0.24*). The model is better than the simple model. 

#Comparing the two models
```{r}
anova(modelA, modelB)
```
The results of the ANOVA test show, that adding **vote_count** and **revenue** variables as predictors to the model significantly improved the model fit (*F = 61.15, p < 0.01*), and that the more complex model explains more variability (*RSS = 139.94*).


#Checking for outliers in modelA
```{r}
cooksdA <- cooks.distance(modelA)
plot(cooksdA, pch = 19, frame = FALSE, main = "Outliers in Model A by Cook's distance")
```

#Checking for outliers in modelB
```{r}
cooksdB <- cooks.distance(modelB)
plot(cooksdB, pch = 19, frame = FALSE, main = "Outliers in Model B by Cook's distance")
```

#Residual diagnostics, assumption checks for the first, simple model (modelA)
##Checking normalitiy
```{r}
qqnorm(residuals(modelA))
qqline(residuals(modelA))
```
##Checking homoscedasticity
```{r}
plot(modelA, which = 1)
```

#Residual diagnostics, assumption checks for the second, multiple model (modelB)
##Checking normalitiy
```{r}
qqnorm(residuals(modelB))
qqline(residuals(modelB))
```

##Checking homoscedasticity
```{r}
plot(modelB, which = 1)
```
##Checking multicollinearity for the multiple regression model
```{r}
vif(modelB)
```
All values are under 3, thus the assumption is not violated.

#Conclusins
The analysis of the dataset about horror movies showed that how people find a movie's quaility and likeability (*idicated by the averahe vote a movie recieves*) is influenced by the budget, but revenue and the number of votes are better predictors of this.