---
title: "Horror Movies"
author: "Fruzsina Hejja"
date: "2024-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(lubridate)
```

In the final project, I will use a dataset containing horror movies and information based on their production and popularity, likeability.

##Reading the dataset
```{r}
horror_movies <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2022/2022-11-01/horror_movies.csv") #Read from the TidyTuesday github page.
```
##Checking the structure of the dataset
```{r}
str(horror_movies)
```
#Exploratory data analysis
Firs a quick overview of all the variables
```{r}
summary(horror_movies)
```
##Creating and exploring the year variable
```{r}
horror_movies <- horror_movies %>%

  mutate(year = year(release_date))
```
###Checking which year most of the movies are from in the dataset
```{r}
horror_movies %>%
  count(year) %>%
  arrange(desc(n))

ggplot(horror_movies, aes(x = year)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "white") +
  labs(title = "Number of horror movies per year",
       x = "Year",
       y = "Number of movies")
  theme_minimal()
```


